The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2015-12-22 DishBeamMeasurements/paper/DishBeamMeasurementsPaper.tex
add new plots to repo and to paper draft

\subsection{Design of the HERA Dish}
%- EM design
%    - departure from large N—small D to economize data volume (mitigate
%      data processing challenges)
%        - easier to identify systematics, to rerun datasets, perform
%          cross checks / jackknives
%    - but this change must be done judiciously, one advantage of simple
%      elements is that they have little or no frequency structure,
%      while a dish+feed system introduces reflections in the time
%      domain, amounting in ripples in the frequency response
%    - how bad are these ripples and can we tune the dish to mitigate
%      their effect in the science?
%    - as discussed above, it is useful to think in the delay domain as
%      the translation from frequency structure introduced by dish
%      reflections and the foreground isolation within the horizon delay
%      is straightforward
%    - the smooth spectral structure of the foregrounds sets a maximum
%      foreground isolation achievable of ~60ns with blackman-harris
%      window
%    - f/D =0.32 maximizes efficiency for a PAPER dipole suspended over
%      the dish
%    - with typical reflection dish/feed reflection coefs, attenuation
%      after 3 round trips will be below 60dB, and 1 and 2 round trips
%      are in wedge, isolated away from EOR window
%    - 14m also falls near the minimum of the cost per performance model

@@

@@

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{aerial.png}
%\caption{Aerial photograph of the two reference dipoles deployed in Galford Meadow at the National Radio Astronomy Observatory--Green Bank 100\,m south of the HERA dish.}
%\label{fig:aerial}
%\end{figure*}

@@

@@

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null1_rel.png}
%\caption{Top: ratio maps of the powers received by two reference dipoles separated by 50\,m, deployed 50\,m south of the HERA dish. Bottom: absolute beam power plots constructed by multiplying the relative power maps by a model beampattern. }
%\label{fig:null1}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null3_rel.png}
%\caption{Same as the previous null experiment except the south reference dipole was moved 5\,m west. The systematics are largely unchanged.}
%\label{fig:null3}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{null4_rel.png}
%\caption{Same as the first null experiment except the reference dipoles are separated by 100\,m and deployed 100\,m south of the HERA dish. The systematics remain at the few percent level near zenith and increase to $\sim10\%$ farther out, with somewhat, but not entirely different angular structure. }
%\label{fig:null4}
%\end{figure*}

@@

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish3_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed lowered 0.5\,m below nominal focus.}
%\label{fig:dish3}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish1_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed at the nominal focus of 4.5\,m.}
%\label{fig:dish1}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish2_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed raised 0.5\,m above nominal focus.}
%\label{fig:dish2}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{dish4_abs_old_ref_model.png}
%\caption{Dish power pattern with the feed raised to its maximum height of 0.8\,m above nominal focus.}
%\label{fig:dish4}
%\end{figure*}

@@

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) for modeled and measured dish beams at 137\,MHz. For measured beams, we calculate the collecting area after interpolation/extrapolation, also with the $60^\circ$ gaussian cutoff (in parentheses).}
\begin{tabular}{| l | l | l |}
\hline
  Airy & 155\,  \\
  CST sim & 67.0\,  \\
  \hline
  Measured, feed at 4\,m & 42.1 (44.3) \\
  Measured, feed at 4.5\,m & 68.5 (73.6) \\ 
  Measured, feed at 5\,m & 77.1 (82.6) \\
  Measured, feed at 5.3\,m & 93.0 (97.9)\\
  \hline
\end{tabular}
\end{table}

@@

We assess the level of foreground leakage into the EOR window due to beam shape, baseline length, and LST. In particular, we built on the study of \citet{nithya15} of the impact of beam shape and horizon response on foreground spectral structure and containment in the wedge.

We simulate delay spectra for several model and measured beams on different baseline lengths for different LSTs and baseline orientations. All these factors affect the level of emission horizon brightening in delay space, the ``pitchfork'' effect, and the level of foreground leakage. We assume a constant beam pattern over that band, with leakage due only to the finite bandwidth and choice of window function. Actual frequency structure imprinted onto the foregrounds by variation of the beam angular pattern or overall gain with frequency (Patra et al.)

We begin by simulating visibilities over a 20\,MHz bandwidth with  400\,kHz resolution centered on 137\,MHz on two different baselines for two model beams (airy, and CST sim) and two measured beams (measured nominal focus beam, and measured nominal focus beam with $60^\circ$ cutoff). We use the short and maximally redundant 14\,m baseline, most useful for delay spectrum analysis, as well as a longer 42\,m baseline, more useful for imaging analysis. From these frequency-sequenced visibilities we calculate the delay spectrum  \citep{perbaselinetechnique} using the Blackman-Harris window following \citet{nithya15}. This window reduces the noise equivalent bandwidth to 10\,MHz, corresponding to $\Delta z\sim0.5$, beyond which evolution of the cosmological signal become significant. Larger bandwidths, though, are useful for foreground estimation and delay-space deconvolution \citep{parsonsandbacker,paper32,paper64}. With this caveat, our simulated delay spectra show the worst case scenario, quantifying the level of foreground isolation without estimation and deconvolution.

Our sky model is comprised of a deep MWA point source survey within 20$^\circ$ of R.A.(J2000) $= 0^\text{h}\,0^\text{m}\,0^\text{s}$ and decl.(J2000) $= -30^\circ\,0'\,0''$ (Carroll et al., in prep.), the shallower but wider MWA commissioning point source survey\citep{MWACS}, the Culgoora catalog\citep{Slee1995}, and the Global Sky Model of Galactic radio emission \citep{gsm}.

Figure \ref{fig:delayspec} shows the delay spectra at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for 14\,m (left pair) and 42\,m (right pair) EW baselines. For reference we plot a 1D model 21\,m power spectrum for $z\sim8$ from \citet{21cmfast}. Consider first the $0^\circ$ LST plots where the Galactic disk has nearly, but not completely set. The effect of horizon response is most visible on the longer 42\,m baseline where the measured beam (red) and CST model (yellow), both show significant brightening at the horizon delays (vertical black lines) due to roughly -30\,dB beam response near the horizon, as discussed by \citet{nithya15,nithya15}. In contrast, the airy model (black) and measured beam with cutoff (blue) show no significant brightening at the horizon, where they appear still to be dominated by the wings of the delay PSF from the emission closer to zero delay.

We also plot (green dashed line) the delay spectrum after subtracting the visibilities computed with the $\theta\sim60^\circ$ cutoff measured beam from those using the full measured beam. This curve may be interpreted as the uncertainty in the ``pitchfork'' horizon brightening due to uncertainty in the beam horizon response, or equivalently as the delay spectrum residual foregrounds after subtracting foregrounds in the well modeled region of the sky (Neben et al. 2015b, submitted). By construction, this subtraction leaves emission from the highest delay regions of the sky (near the horizon on opposite sides of the baseline), but also leaves some low delay emission on opposite sides of a line perpendicular to the baseline. However, this additional low delay emission near the horizon gets drowned out by that from near zenith where the beam is substantially stronger.

On the shorter 14\,m baseline, the horizon brightening is smaller because there is roughly an order of magnitude more emission at zero delay, whose thereby stronger delay wings, which are also relatively wider compared to the narrower foreground wedge, cover up much of the edge brightening. At $60^\circ$ LST, effect of horizon response is much smaller as the Galactic disk is now entirely below the horizon.

The level of leakage depends also on baseline orientation which sets which directions on the sky map to large delays. Figure \ref{fig:delayspec2} shows the delay spectra at $0^\circ$ LST on 14\,m NE (top pair) and SE (bottom pair) baselines. Both show increased edge brightening compared to the 14\,m EW baseline. \citet{nithya15} proposes to identify through modeling the baselines most susceptible to this edge brightening at any given time, and simply exclude them from the power spectrum analysis.

%\begin{figure*}[h]
%\includegraphics[width=6.7in]{LST0deg_14m_42m_EWbaselines_dish1_blackman.png}
%\includegraphics[width=6.7in]{LST60deg_14m_42m_EWbaselines_dish1_blackman.png}
%\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
%\label{fig:delayspec}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.7in]{LST0deg_14m_42m_NEslantbaselines_dish1_blackman.png}
%\includegraphics[width=6.7in]{LST0deg_14m_42m_SEslantbaselines_dish1_blackman.png}
%\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ LST for NE (top pair) and SE (bottom pair) baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
%\label{fig:delayspec2}
%\end{figure*}

>>>>

\subsection{Design of the HERA Dish}

\begin{figure}[h]
\includegraphics[width=3.4in]{IMG_0353.JPG}
\caption{Aerial photograph of the two reference dipoles deployed in Galford Meadow at the National Radio Astronomy Observatory--Green Bank 100\,m south of the HERA dish.}
\label{fig:dishphoto}
\end{figure}

\begin{figure}[h]
\includegraphics[width=3.4in]{feed.jpg}
\caption{Aerial photograph of the two reference dipoles deployed in Galford Meadow at the National Radio Astronomy Observatory--Green Bank 100\,m south of the HERA dish.}
\label{fig:dishphoto}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=7in]{dave197_rich195_airy_beams.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec}
\end{figure*}

@@

@@

\begin{figure*}[h]
\centering
\includegraphics[width=6.7in]{null_expt_rel_beam_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:nullexptplots}
\end{figure*}

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.7in]{measured_beams_and_models_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeammaps}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=6.7in]{measured_beams_and_models_slices.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeamslices}
\end{figure*}

@@

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) of measured 137\,MHz beams and corresponding power spectrum SNR for HERA-127.}
\begin{tabular}{| l | l | l |}
\hline
Beam & $A_\text{eff}$ (m$^2$) & SNR (pess, mod, opt)\\
\hline
  Airy pattern & 155 & 9.1, 11.0, 37.2  \\
  %CST sim & 67.0 &  \\
  %Measured, feed at 4\,m & 44.3 &  \\
    Measured, feed at 5.3\,m & 97.9 & 8.2, 8.3, 29.2 \\
    Measured, feed at 5\,m & 82.6 & 6.8, 6.9, 26.5 \\
    Measured, feed at 4.5\,m & 73.6 & 6.4, 6.5, 24.8 \\ 
  \hline
\end{tabular}
\end{table}

@@

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_manyLSTs.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecmanyLSTs}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{gsm_kelvin_LST_2deg_and_62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:gsmplots}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_LST2deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecbadLST}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_LST62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecgoodLST}
\end{figure*}
